---
title: "Plain-language decision rulings: estimand-aware multiverse meta-analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plain-language decision rulings: estimand-aware multiverse meta-analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7)
```

```{r setup}
library(metaMultiverse)
```

This vignette demonstrates the estimand-aware multiverse workflow on
simulated data: plain-language fork rulings (with E/U/N/X letter
aliases), per-domain configs with
warrants, load-time validation, the crossed estimand grid, execution with
full traceability, plausibility flagging, fragility metrics, and reporting.

## Decision rulings

Every analytic or clinical decision ("fork") gets an individual
plain-language ruling (letter aliases from the source framework are
accepted everywhere):

```{r crosswalk}
knitr::kable(decision_crosswalk())
```

- **equivalent** (alias E): defensible alternatives expected to be interchangeable.
  Varied within the multiverse; movement beyond a declared
  equivalence margin is
  flagged for reclassification.
- **uncertain** (alias U): defensible alternatives that plausibly matter, but the
  literature cannot yet rank them. Varied within the multiverse and tracked
  as metadata.
- **changes_question** (alias N): levels answer *different clinical questions*.
  Crossing question-changing forks defines the **estimand cells** of the
  grid; effect
  summaries are never pooled across cells.
- **indefensible** (alias X): historically used but not defensible; excluded from inference,
  quantified separately in the bias-decomposition layer.

This adapts Del Giudice & Gangestad's (2021) framework, in which N-type
decisions are fixed to the best option and U-type decisions split into
separate multiverses. Here the split criterion is the *estimand*: N forks
split because they change the clinical question; U forks vary within
because they do not. The reclassification pathway (groundwork moves U
forks to fixed decisions, their inferior alternatives to X) is the same.

## A simulated domain

```{r simulate}
dat <- simulate_multiverse_data(n_studies = 40, es_per_study = 1:2, seed = 7)
head(dat)
```

Flag implausible effect sizes (Harrer et al. 2025). The reference
distribution is an explicit choice; missing risk-of-bias ratings count as
high risk by project convention:

```{r flags}
dat <- add_plausibility_flags(dat, high_rob = dat$rob == 1, reference = "all")
table(dat$flag_profile)
```

## Config with warrants

```{r config}
cfg <- domain_config(
  domain = "simulated_demo",
  forks = list(
    fork("comparator", type = "data", decision = "changes_question",
         family = "clinical",
         column = "comparator",
         levels = list(wl = "wl", cau = "cau",
                       all_controls = c("wl", "cau", "other")),
         reference = "all_controls",
         justification = list(
           basis = "citation",
           text = "Control condition changes the clinical question",
           citation = "Cuijpers et al. 2019")),
    plausibility_fork(),
    fork("ma_method", type = "analysis", decision = "uncertain",
         family = "analytic",
         levels = c("reml", "pm"), reference = "reml",
         justification = "Both defensible tau2 estimators; ranking unresolved"),
    fork("dependency", type = "analysis", decision = "uncertain",
         family = "analytic",
         levels = c("aggregate", "modeled"), reference = "aggregate",
         justification = "Both defensible dependency strategies"),
    fork("model_type", type = "analysis", decision = "indefensible",
         family = "analytic",
         levels = c("random_effects", "fixed_effect"),
         reference = "random_effects",
         overrides = list(fixed_effect = list(ma_method = "fe")),
         justification = list(
           basis = "citation",
           text = "Fixed-effect under clear heterogeneity is indefensible;
                   quantified in the bias layer only",
           citation = "Borenstein et al. 2010"))
  ),
  estimand = list(population = "simulated adults",
                  intervention = "simulated psychotherapy",
                  outcome = "symptoms", timepoint = "post-treatment"),
  fixed_decisions = list(
    effect_metric = list(
      value = "Hedges' g",
      justification = list(basis = "citation",
                           text = "Standard small-sample-corrected SMD",
                           citation = "Hedges 1981"))
  )
)
cfg
```

The warrant table is generated automatically -- no fork exists without a
machine-readable justification:

```{r warrants}
warrant_table(cfg)[, c("element", "name", "decision", "family", "basis")]
```

A preregisterable, time-stamped export of configs + warrants + estimand
declarations is available via `export_prereg_config(cfg, "prereg.qmd")`.

## Validation and the availability matrix

Configs are validated against the data at load time; absent levels are
informative errors, small strata are set aside as too_few_studies
(min_studies = 5, strict rerun at 10), and levels with identical realized subsets are marked as
duplicates rather than pretending to vary:

```{r validate}
v <- validate_domain_config(cfg, dat)
v$fork_status
```

## The crossed estimand grid

One grid per domain: question-changing fork levels define estimand
cells; equivalent/uncertain and analysis forks vary within cells;
indefensible forks run only in the bias layer against the
reference specification:

```{r grid}
grid <- build_specification_grid(cfg, dat)
summary(grid)
```

## Execution with full traceability

```{r run}
res <- run_multiverse_grid(grid, rho = 0.6, progress = FALSE)
attrition_summary(res)
```

Every estimate is traceable: `lookup_spec(res, spec_id)` returns the spec
definition, estimand declaration, fork warrants, result, audit flags, and
attrition entries.

## Effect summaries respect the two-layer rule

```{r summaries}
summarize_effects(res)
```

Pooling effects across estimand cells is refused at runtime:

```{r guard, error = TRUE}
assert_single_estimand(res$results, "a pooled median")
```

Divergence between cells is its own labeled quantity:

```{r divergence}
str(summarize_divergence(res), max.level = 1)
```

## Fragility metrics (within a fixed clinical question)

```{r fragility}
fragility_metrics(res, cell = "comparator=cau")
```

M2 (specification-I^2) is the provisional primary metric: it expresses
analytic divergence relative to sampling uncertainty, so "fragile" cannot
be an artifact of a small, noisy database. Equivalent-ruled forks that
move estimates
beyond their equivalence margin -- and any fork dominating the
variance -- are
escalated for reclassification:

```{r tolerances}
check_fork_tolerances(res, cell = "comparator=cau")
```

## Report

```{r report, results = "hide"}
report <- report_grid_result(res)
```

The report leads with the justified specification count, labels all
frequencies as descriptive, includes the worst-universe audit, the bias
layer, the attrition table, and the warrant table.

## Two-layer uncertainty visualization

```{r plots, fig.height = 4.5}
plot_two_layer(res, "comparator=cau")
plot_grid_spec_curve(res)
```
